---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

## Overview
Simple error detection and version control for the importing, exporting, and transfering of data using CRC32 checksum. A hash file is generated for each dataset containing the summary hashes for all columns and rows along with col/row names. Each file is created before exporting a dataset and compared to hashes after importing to ensure consistency in the datasets between working sessions of each user. In addition, each file contains hashes organized by row and column which can be used for easy debugging of potential errors.

Tutorial/Documentation:

* Host: http://j-huang.com/resources/tutorial/vckii
* Open Science Framework(OSF): https://osf.io/xgd7q/

## 1. Installing package
First, install the R package “vckii”. The package can also be download from the following repositories:

* Host: http://j-huang.com/resources/pk/vckii.tar.gz
* Github: https://github.com/jhudev/vckii
* Open Science Framework(OSF): https://osf.io/xgd7q/

```{r eval = FALSE}
#Install R package from host site
install.packages("http://j-huang.com/resources/pk/vckii.tar.gz")

# Or the development version from GitHub
# install.packages("devtools")
devtools::install_github("jhudev/vckii")

# Source: https://osf.io/xgd7q/
```

## 2. Generating a hash file and data verification
The two primary functions are makeHash() and checkHash(). 

Exporting: The makeHash() function is used to generate a hash file and should be used during the data exporting process. The hash file is a simple text file that contains a set of hashes used for data verification and debugging any issues during the importing process. The structure of the file is detailed in the following section (Section 3). Therefore, this file should be provided along with the exported data for verification by the receiver. 

Importing: The checkHash() function is used to verify a dataset after importing into the working environment by comparing it with its corresponding hash file. It will return either TRUE, if all checksums match those in the hash file; otherwise, it will return FALSE. If FALSE, the hash file can be used to determine the source of any inconsistencies and assist in the debugging of the importing process.

A simplified example of the process is provided below.

```{r eval = FALSE}
#Step 1: User 1 exporting dataset “iris”
makeHash(iris, “iris_hash.txt”)
write.csv(iris, “iris.csv”)

#Step 2: Transferring data and hash files from User 1 to User 2

#Step 3: User 2 importing dataset “iris.csv” and validating with “iris_hash.txt”
iris_2 <- read.csv(“iris.csv”)
checkHash(iris_2, “iris_hash.txt”)
#Output >> TRUE
```

